---
title: 'Exercise 9: Analysis'
author: "Ethan vanderWilden"
date: "11/12/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review

```{r packages, include = FALSE}
library("here")
library("haven")
library("Rmisc")
library("magrittr")
library("tidyverse")
library("ggplot2")
library("broom")
#library("MASS")
library("estimatr")
library("modelsummary")
library("stargazer")
library("naniar")
```

3. Set up your files and folder structure.
```{r}
getwd()
#already in the right place because I am in my exercises R proj
```

4. Read the ANES .dta data into R using the `here` package.
```{r}
anes2016 <- read_dta("anes_timeseries_2016.dta") #loading in data
```

5. Download the ANES 2016 codebook (available on the `ps811/data` repository). We will look at the full sample variables.

Done!

6. You want to know whether owning a house (pre-election) affects which party the respondent choose to contribute to (post-election). Identify these variables from the codebook and rename the variables to names that are easier to reference.

```{r}
#table(anes2016$V161334) #home ownership variable
anes2016$homeowner <- anes2016$V161334 #variable now stored under "homeowner"
anes2016 <- mutate(anes2016,
                   homeowner = case_when(
                     homeowner == 1 ~ 0, #non-homeowner (paying rent)
                     homeowner == 4 ~ 0, #non-homeowner (other arrangement)
                     homeowner == 2 ~ 1, #homeowner (paying mortgage)
                     homeowner == 3 ~ 1)) #homeowner (no payments due)


#table(anes2016$V162014a) #which party respondent donated to
anes2016$party_donate <- anes2016$V162014a
anes2016 <- mutate(anes2016,
                   party_donate = case_when(
                     party_donate == 1 ~ 0, #Donates democrat
                     party_donate == 2 ~ 1)) #Donates republican
#NOTE: not including donations to both parties or to other parties because I think that deviates from the question at hand
```

7. Now identify pre-election demographic variables, such as age, gender, and race. Manipulate these variables in ways that you believe would best capture these demographics and explain why you manipulated these variables that way you did. Rename these variables to names that are easier to reference.

```{r}
#table(anes2016$V161267) #age
anes2016$age <- anes2016$V161267
anes2016$age <- na_if(anes2016$age, -9) #getting rid of NAs
anes2016$age <- na_if(anes2016$age, -8)

#table(anes2016$V161342) #gender
anes2016$gender <- anes2016$V161342
anes2016 <- mutate(anes2016,
                   gender = case_when(
                     gender == 1 ~ 0, #gender = MALE
                     gender == 2 ~ 1)) #gender = FEMALE


#EXPLANATION: This is definitely problematic to be excluding the data to only 2 binary
#genders. However, I'm not too sure of how we might run a regression on this data if we
#cannot make the gender variable numeric, and with the inclusion of more than two genders,
#I think the numeric nature of this data would no longer make sense. I recognize that if I
#were to actually be writing this paper, I would look more into how to treat this data
#without being exclusionary.

#table(anes2016$V161310x) #race
anes2016$race <- anes2016$V161310x
anes2016 <- mutate(anes2016,
                   race = case_when(
                     race == 1 ~ 0, #race = WHITE (non-Hispanic)
                     race == 2 ~ 1, #race = BLACK (non-Hispanic)
                     race == 3 ~ 1, #race = ASIAN/HAWAIIAN/PAC ISLANDER
                     race == 4 ~ 1, #race = NATIVE AMERICAN
                     race == 5 ~ 1, #race = HISPANIC
                     race == 6 ~ 1)) #race = OTHER
#Race variable as binary (0 = White non-Hispanic, 1 = Non-White)

#EXPLANATION: Again, I recognize that it is problematic to turn race into a binary.
#However, I think that if I want to run a regression, I need all of the variable to make
#sense as numeric. Including all of the mentioned races would make the order of the data
#make no sense.

```


8. Provide descriptive summaries for each variable.

```{r}
anes_test <- anes2016[1843:1847]
anes_test <- na.omit(anes_test)
datasummary((`Homeownership` = homeowner) + (`Party Donation` = party_donate) +
              (`Age` = age) + (`Gender` =gender) + (`Race` = race)
            ~ Mean + SD + Min + Max,
            data = anes_test,
            output = 'latex')

```



\newpage
9. Run an appropriate regression analysis and insert the table into the R Markdown document.

```{r , results='asis'}
model1 <- glm(anes_test$party_donate ~ anes_test$homeowner)
model2 <- glm(anes_test$party_donate ~ anes_test$homeowner + anes_test$age +
                        anes_test$gender + anes_test$race)

# Copying code from https://cimentadaj.github.io/blog/2016-08-22-producing-stargazer-tables-with-odds-ratios-and-standard-errors-in-r/producing-stargazer-tables-with-odds-ratios-and-standard-errors-in-r/


stargazer2 <- function(model, odd.ratio = F, ...) {
  if(!("list" %in% class(model))) model <- list(model)
    
  if (odd.ratio) {
    coefOR2 <- lapply(model, function(x) exp(coef(x)))
    seOR2 <- lapply(model, function(x) exp(coef(x)) * summary(x)$coef[, 2])
    p2 <- lapply(model, function(x) summary(x)$coefficients[, 4])
    stargazer(model, coef = coefOR2, se = seOR2, p = p2, ...)
    
  } else {
    stargazer(model, ...)
  }
}

#Making table with odds ratios for logit regression
models <- list(model1, model2)
stargazer2(models, odd.ratio = T, type = "latex",
           title = "Odds Ratios: Effect of Homeownership on Partisan Donation",
          covariate.labels = c("Homeownership", "Age", "Gender", "Race"),
          dep.var.labels   = "Party Donated To",
          header = FALSE)

```

10. Create a coefficient plot based on the above table.

```{r}
#Coefficient plot
vars <- c('anes_test$homeowner' = 'Homeowner?',
          'anes_test$age' = 'Age',
          'anes_test$gender' = 'Gender',
          'anes_test$race' = 'Race')
modelplot(model2, coef_map = vars)

```

\newpage 
## Your project (Tables above included)

Now it's your turn. Use the tools you used today to conduct data analysis for one of your final seminar papers.

1. Create a descriptive statistics summary table for your main variables of interest. Note the number of observations.

```{r Cleaning Data, include=FALSE}
library(haven)
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)

barometer <- read_sav("3269.sav") #load in data

#subset to get the relevant variables to my study
mydata <- select(barometer, C12,
                 C12A,B22, C3, C9, C10, C4_1,C4_2,C4_3,C4_4,C4_6)

#rename the variables so I have an easier time navigating the data
mydata <- mydata %>%
  rename(religion = C12,
         rel_attendance =C12A,
         partyvote = B22,
         voter_ideology = C3,
         sex = C9,
         age = C10,
         psoe_ideology = C4_1,
         pp_ideology = C4_2,
         vox_ideology = C4_3,
         podemos_ideology = C4_4,
         cs_ideology = C4_6)

    ####For Each Variable, I want to make NAs uniform (in the dataset, they are 
          #often as numbers and differ by variable)

#IDEOLOGY
table(mydata$voter_ideology)
mydata <- mutate(mydata,
                 voter_ideology = case_when(
                   voter_ideology == 1~1, voter_ideology == 2~2,
                   voter_ideology == 3~3, voter_ideology == 4~4,
                   voter_ideology == 5~5, voter_ideology == 6~6,
                   voter_ideology == 7~7, voter_ideology == 8~8,
                   voter_ideology == 9~9, voter_ideology == 10~10))

#SIMPLIFIED IDEOLOGY
mydata$simple_ideology <- mydata$voter_ideology
mydata <- mutate(mydata,
                 simple_ideology = case_when(
                   simple_ideology == 1~0, voter_ideology == 2~0,
                   voter_ideology == 3~0, voter_ideology == 4~0,
                   voter_ideology == 5~1, voter_ideology == 6~1,
                   voter_ideology == 7~2, voter_ideology == 8~2,
                   voter_ideology == 9~2, voter_ideology == 10~2))

#RELIGION
table(mydata$religion)
mydata <- mutate(mydata,
                 religion = case_when(
                   religion == 1~ "Practicing Catholic", 
                   religion == 2~ "Non-Practicing Catholic",
                   religion == 3~ "Other Religion",
                   religion == 4~ "Agnostic/Indifferent/Athiest"))

#RELIGIOUS ATTENDANCE
table(mydata$rel_attendance)
mydata <- mutate(mydata,
                 rel_attendance = case_when(
                   rel_attendance == 1~1, rel_attendance == 2~2,
                   rel_attendance == 3~3, rel_attendance == 4~4,
                   rel_attendance == 5~5, rel_attendance == 6~6))

#PARTY VOTE
mydata <- mutate(mydata,
                      partyvote = case_when(
                        partyvote == 1 ~ "PP", partyvote == 2 ~ "PSOE",
                        partyvote == 4 ~ "Cs",partyvote == 21 ~ "Podemos",
                        partyvote == 18 ~ "Vox"))

#SEX
table(mydata$sex) 
mydata <- mutate(mydata,
                 sex = case_when(
                   sex == 1~0, sex == 2~1)) # MALE = 0, FEMALE = 1

#AGE
table(mydata$age)

table(mydata$religion, mydata$partyvote)
prop.table(table(mydata$religion, mydata$simple_ideology), margin = 2)
```


```{r Numeric Variables Table}
datasummary((`Religious Attendance` = na.omit(rel_attendance)) + 
              (`Voter Ideology` = na.omit(voter_ideology)) +
              (`Simplified Ideology` = na.omit(simple_ideology)) +
              (`Sex` = na.omit(sex)) + (`Age` = na.omit(age))
            ~ (`N` = length) + Mean + SD + Min + Max,
            data = mydata,
            output = 'latex',
            title = "Descriptive Statistics for Numeric Variables")
```

```{r Religion Variable}
#Descriptive Stats for Categorical Religion Variable
datasummary_skim(mydata$religion, type='categorical', 
                 title = "Religion", output = 'latex')
```

```{r Party Vote Variable}
#Descriptive Stats for Categorical Party Vote Variable
datasummary_skim(mydata$partyvote, type='categorical', 
                 title = "Party Supported", output = 'latex')

```






2. If you are planning to run a regression, please write out the regression formula. Please take into consideration the dependent variable and its distribution. If you already have the data, you may go ahead and run it. If you do not have the data and is in the process of collecting it, write out the formula. Pre-analysis plans are becoming more common in the discipline, so being able to record what you *plan* to do is becoming increasingly more important.



I am still working out exactly what tests I am going to do. I think I will do a difference in proportions test to demonstrate that the radical right party does not dominate the 'religious vote' among parties on the right.

```{r ,include=FALSE, echo=FALSE}
prop.table(table(mydata$religion, mydata$partyvote), margin = 2)
```


I am also going to do a regression to assess the relationship between religious attendance and ideology. I still definitely need to do more thinking on what exactly I want to do with my statistics. However, I could run a regression to assess this relationship:

\begin{center}
$RELIGIOSITY_i = \beta_0 + \beta_1 IDEOLOGY_i + \beta_2 AGE_i + \beta_3 SEX_i + \epsilon_i$
\end{center}

```{r, results='asis'}
no_control <- lm(mydata$voter_ideology ~ mydata$rel_attendance)
model <- lm(mydata$voter_ideology ~ mydata$rel_attendance + mydata$age + mydata$sex)

stargazer(no_control, model,  type = "latex",
           title = "Effect of Religious Attendance  on Voter Ideology",
          covariate.labels = c("Religious Attendance Level", "Age", "Sex"),
          dep.var.labels   = "Voter Ideology",
          header = FALSE)

```



# Submit

Email me (mshieh2@wisc.edu) the link to your `ps811-exercises` repository when you are done.